---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. If you can't, peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
library(rethinking)

#Make the grid approximation
dens = 10000
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = rep(1, dens) #start with implementing a flat prior
likelihood = dbinom(3, size = 6, prob = p_grid)
unstd.posterior = likelihood*prior
posterior = unstd.posterior/sum(unstd.posterior)


#Calculate the likelihood that riccardo knows more than chance
sum(posterior[p_grid > 0.5])
#50 % chance of performing above chance level

library(ggplot2)
data = data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

#Plot of Riccardos CogSci Knowledge
RF = ggplot(data, aes(grid, posterior))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Riccardo")+
  ylim(0,0.0015)

#Use the quadratic approximation to find Riccardo's knowledge
#assumes that the posterior will be a gaussian distribution -> And all you need to describe this is the mean and the SD
Riccardo.qa = map(
  alist(
    c ~ dbinom(6,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 3))

precis(Riccardo.qa)


samplesRF = sample(p_grid, size = 1e4, replace = T, prob = posterior)

#Use HPDI 
HPDI(samplesRF, prob = 0.5)



```



2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.

2a. Produce plots of the prior, and posterior for each teacher.

```{r}



#---------------------------------------------#
#Kristian
#Make the grid approximation
likelihood = dbinom(2, size = 2, prob = p_grid)
unstd.posterior = likelihood*prior
KT_posterior = unstd.posterior/sum(unstd.posterior)

KT_data = data.frame(grid = p_grid, posterior = KT_posterior, prior = prior, likelihood = likelihood)

#Plot of KTs CogSci Knowledge
KT = ggplot(KT_data, aes(grid, KT_posterior))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Kristian")+
  ylim(0,0.0015)

samplesKT = sample(p_grid, size = 1e4, replace = T, prob = KT_posterior)

#Use HPDI 
HPDI(samplesKT, prob = 0.5)



#Doesn't work...
tylen.qa = map(
  alist(
    c ~ dbinom(2,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 2))

precis(tylen.qa)




#---------------------------------------------#
#JOSH <3  (we miss Josh)
#Make the grid approximation
likelihood = dbinom(160, size = 198, prob = p_grid)
unstd.posterior = likelihood*prior
JOSH_posterior = unstd.posterior/sum(unstd.posterior)

JOSH_data = data.frame(grid = p_grid, posterior = JOSH_posterior, prior = prior, likelihood = likelihood)

#Plot of Joshs CogSci Knowledge
Josh = ggplot(JOSH_data, aes(grid, JOSH_posterior))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Josh")+
  ylim(0,0.0015)

samplesJOSH = sample(p_grid, size = 1e4, replace = T, prob = JOSH_posterior)

#Use HPDI 
HPDI(samplesJOSH, prob = 0.5)


josh.qa = map(
  alist(
    c ~ dbinom(198,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 160))

precis(josh.qa)


#---------------------------------------------#
#MW
#Make the grid approximation
likelihood = dbinom(66, size = 132, prob = p_grid)
unstd.posterior = likelihood*prior
MW_posterior = unstd.posterior/sum(unstd.posterior)

MW_data = data.frame(grid = p_grid, posterior = MW_posterior, prior = prior, likelihood = likelihood)

#Plot of KTs CogSci Knowledge
MW = ggplot(MW_data, aes(grid, MW_posterior))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Mikkel")+
  ylim(0, 0.0015)

samplesMW = sample(p_grid, size = 1e4, replace = T, prob = MW_posterior)

#Use HPDI 
HPDI(samplesMW, prob = 0.5)

mikkel.qa = map(
  alist(
    c ~ dbinom(132,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 66))

precis(mikkel.qa)



#Arrange the plots together
gridExtra::grid.arrange(RF, KT, Josh, MW)

```




3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.
```{r}
#Make the grid approximation
dens = 10000
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = dnorm(p_grid, mean = 0.8, sd = 0.2) #Now we need to implement a new prior
likelihood = dbinom(3, size = 6, prob = p_grid)
unstd.posterior = likelihood*prior
posterior = unstd.posterior/sum(unstd.posterior)

#plots
RF2_data = data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

#Plot of Riccardos CogSci Knowledge
RF2 = ggplot(RF2_data, aes(grid, posterior))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Riccardo")+
  ylim(0, 0.0015)

RF2.qa = map(
  alist(
    c ~ dbinom(6,p),
    p ~ dnorm(mean = 0.8, sd = 0.2) 
  ),
  data = list(c = 3))

precis(RF2.qa)


#---------------------------------------------#
#Kristian
#Make the grid approximation
likelihood = dbinom(2, size = 2, prob = p_grid)
unstd.posterior = likelihood*prior
KT_posterior2 = unstd.posterior/sum(unstd.posterior)

KT_data2 = data.frame(grid = p_grid, posterior = KT_posterior2, prior = prior, likelihood = likelihood)

#Plot of KTs CogSci Knowledge
KT2 = ggplot(KT_data2, aes(grid, KT_posterior2))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Kristian")+
  ylim(0, 0.0015)

samplesKT2 = sample(p_grid, size = 1e4, replace = T, prob = KT_posterior2)

mean(samplesKT2)
sd(samplesKT2)
PI(samplesKT2, prob = 0.89)


#---------------------------------------------#
#JOSH <3  (we miss Josh)
#Make the grid approximation
likelihood = dbinom(160, size = 198, prob = p_grid)
unstd.posterior = likelihood*prior
JOSH_posterior2 = unstd.posterior/sum(unstd.posterior)

JOSH_data2 = data.frame(grid = p_grid, posterior = JOSH_posterior2, prior = prior, likelihood = likelihood)

#Plot of Joshs CogSci Knowledge
Josh2 = ggplot(JOSH_data2, aes(grid, JOSH_posterior2))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Josh")+
  ylim(0, 0.0015)

Josh2.qa = map(
  alist(
    c ~ dbinom(198,p),
    p ~ dnorm(mean = 0.8, sd = 0.2) #Uniform prior
  ),
  data = list(c = 160))

precis(Josh2.qa)

#---------------------------------------------#
#MW
#Make the grid approximation
likelihood = dbinom(66, size = 132, prob = p_grid)
unstd.posterior = likelihood*prior
MW_posterior2 = unstd.posterior/sum(unstd.posterior)

MW_data2 = data.frame(grid = p_grid, posterior = MW_posterior2, prior = prior, likelihood = likelihood)

#Plot of MWs CogSci Knowledge
MW2 = ggplot(MW_data2, aes(grid, MW_posterior2))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Mikkel")+
  ylim(0, 0.0015)


MW2.qa = map(
  alist(
    c ~ dbinom(132,p),
    p ~ dnorm(mean = 0.8, sd = 0.2) #Uniform prior
  ),
  data = list(c = 66))

precis(MW2.qa)

#Arrange the plots together
gridExtra::grid.arrange(RF2, KT2, Josh2, MW2)



```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
#100 more datapoints with a flat prior
#Make the grid approximation
dens = 10000
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = rep(1, dens) #flat prior
likelihood = dbinom(300, size = 600, prob = p_grid)
unstd.posterior = likelihood*prior
RFposterior3 = unstd.posterior/sum(unstd.posterior)

#plots
RF3_data = data.frame(grid = p_grid, posterior = RFposterior3, prior = prior, likelihood = likelihood)

#Plot of Riccardos CogSci Knowledge
RF3 = ggplot(RF3_data, aes(grid, RFposterior3))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Riccardo")+
  ylim(0,0.02)

RF3.qa = map(
  alist(
    c ~ dbinom(600,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 300))

precis(RF4.qa)


#---------------------------------------------#
#Kristian
#Make the grid approximation
likelihood = dbinom(200, size = 200, prob = p_grid)
unstd.posterior = likelihood*prior
KT_posterior3 = unstd.posterior/sum(unstd.posterior)

KT_data3 = data.frame(grid = p_grid, posterior = KT_posterior3, prior = prior, likelihood = likelihood)

#Plot of KTs CogSci Knowledge
KT3 = ggplot(KT_data3, aes(grid, KT_posterior3))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Kristian")+
  ylim(0,0.02)

samplesKT3 = sample(p_grid, size = 1e4, replace = T, prob = KT_posterior3)

mean(samplesKT3)
sd(samplesKT3)
PI(samplesKT3, prob = 0.89)

#---------------------------------------------#
#JOSH <3  (we miss Josh)
#Make the grid approximation
likelihood = dbinom(16000, size = 19800, prob = p_grid)
unstd.posterior = likelihood*prior
JOSH_posterior3 = unstd.posterior/sum(unstd.posterior)

JOSH_data3 = data.frame(grid = p_grid, posterior = JOSH_posterior3, prior = prior, likelihood = likelihood)

#Plot of Joshs CogSci Knowledge
Josh3 = ggplot(JOSH_data3, aes(grid, JOSH_posterior3))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Josh")+
  ylim(0,0.02)

Josh3.qa = map(
  alist(
    c ~ dbinom(19800,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 16000))

precis(Josh3.qa)

#---------------------------------------------#
#MW
#Make the grid approximation
likelihood = dbinom(6600, size = 13200, prob = p_grid)
unstd.posterior = likelihood*prior
MW_posterior3 = unstd.posterior/sum(unstd.posterior)

MW_data3 = data.frame(grid = p_grid, posterior = MW_posterior3, prior = prior, likelihood = likelihood)

#Plot of MWs CogSci Knowledge
MW3 = ggplot(MW_data3, aes(grid, MW_posterior3))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Mikkel")+
  ylim(0,0.02)

mikkel3.qa = map(
  alist(
    c ~ dbinom(13200,p),
    p ~ dunif(0,1) #Uniform prior
  ),
  data = list(c = 6600))

precis(mikkel3.qa)

#Arrange the plots together
gridExtra::grid.arrange(RF3, KT3, Josh3, MW3)


#100 more datapoints with a optimistic prior

#Make the grid approximation
dens = 10000
p_grid = seq(from = 0, to = 1, length.out = dens)
prior = dnorm(p_grid, mean = 0.8, sd = 0.2) #Now we need to implement a new prior
likelihood = dbinom(300, size = 600, prob = p_grid)
unstd.posterior = likelihood*prior
RFposterior4 = unstd.posterior/sum(unstd.posterior)

#plots
RF4_data = data.frame(grid = p_grid, posterior = RFposterior4, prior = prior, likelihood = likelihood)

#Plot of Riccardos CogSci Knowledge
RF4 = ggplot(RF4_data, aes(grid, RFposterior4))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Riccardo")+
  ylim(0,0.02)

RF4.qa = map(
  alist(
    c ~ dbinom(600,p),
    p ~ dnorm(mean = 0.8, sd = 0.2) 
  ),
  data = list(c = 300))

precis(RF4.qa)


#---------------------------------------------#
#Kristian
#Make the grid approximation
likelihood = dbinom(200, size = 200, prob = p_grid)
unstd.posterior = likelihood*prior
KT_posterior4 = unstd.posterior/sum(unstd.posterior)

KT_data4 = data.frame(grid = p_grid, posterior = KT_posterior4, prior = prior, likelihood = likelihood)

#Plot of KTs CogSci Knowledge
KT4 = ggplot(KT_data4, aes(grid, KT_posterior4))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Kristian")+
  ylim(0,0.02)


samplesKT4 = sample(p_grid, size = 1e4, replace = T, prob = KT_posterior4)

mean(samplesKT4)
sd(samplesKT4)
PI(samplesKT4, prob = 0.89)

#---------------------------------------------#
#JOSH <3  (we miss Josh)
#Make the grid approximation
likelihood = dbinom(16000, size = 19800, prob = p_grid)
unstd.posterior = likelihood*prior
JOSH_posterior4 = unstd.posterior/sum(unstd.posterior)

JOSH_data4 = data.frame(grid = p_grid, posterior = JOSH_posterior4, prior = prior, likelihood = likelihood)

Josh4.qa = map(
  alist(
    c ~ dbinom(19800,p),
    p ~ dnorm(mean = 0.8, sd = 0.2) 
  ),
  data = list(c = 16000))

precis(Josh4.qa)

#Plot of Joshs CogSci Knowledge
Josh4 = ggplot(JOSH_data4, aes(grid, JOSH_posterior4))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Josh")+
  ylim(0,0.02)

#---------------------------------------------#
#MW
#Make the grid approximation
likelihood = dbinom(6600, size = 13200, prob = p_grid)
unstd.posterior = likelihood*prior
MW_posterior4 = unstd.posterior/sum(unstd.posterior)

MW_data4 = data.frame(grid = p_grid, posterior = MW_posterior4, prior = prior, likelihood = likelihood)

#Plot of MWs CogSci Knowledge
MW4 = ggplot(MW_data4, aes(grid, MW_posterior4))+
  geom_point()+
  geom_line()+
  theme_classic()+
  geom_line(aes(grid, prior/dens), color = 'red')+
  xlab("Knowledge of CogSci")+
  ylab("Posterior probability")+
  ggtitle("Mikkel")+
  ylim(0,0.02)

MW4.qa = map(
  alist(
    c ~ dbinom(13200,p),
    p ~ dnorm(mean = 0.8, sd = 0.2) 
  ),
  data = list(c = 6600))

precis(MW4.qa)

#Arrange the plots together
gridExtra::grid.arrange(RF4, KT4, Josh4, MW4)


```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
###ASK! - have a prior of 0.5, 
no, we have a normal distribution centered around 0.5

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



